---
title: "INTERVAL Misexpression Fig.4 and associated supplementary Figs"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr) 
library(RColorBrewer)
library(ggrastr)
```

```{r}
ticks_size <- 0.75
border_size <- 1.5

sv_vrnt_colors <- c('DEL' = '#E04B4B', 
                    'DUP' ='#00C750', 
                    "Non-carrier" = "grey"
                    )
```


```{r}
# output directory
wkdir <- "/lustre/scratch126/humgen/projects/interval_rna/interval_rna_seq/thomasVDS/misexpression_v3"
# Fig. 2
fig4_out_dir <- file.path(wkdir, "figs/fig_4")
dir.create(file.path(fig4_out_dir))
# Supplementary Fig 13
supp_fig13_out_dir <- file.path(wkdir, "figs/supp_fig_13")
dir.create(file.path(supp_fig13_out_dir))
```

```{r}
tx_readthrough_path <- file.path(wkdir, '6_misexp_dissect/tx_readthrough/combine/del_dup_readthrough_region_wgs.tsv')
tx_readthrough_df <- read.csv(tx_readthrough_path, sep="\t")

tx_readthrough_carrier_df <- tx_readthrough_df[tx_readthrough_df$carrier == "DEL" | tx_readthrough_df$carrier == "DUP", ]
tx_readthrough_noncarrier_df <- tx_readthrough_df[tx_readthrough_df$carrier == "Non-carrier", ]
```

### FBNC vs FPKM z-score carrier and non-carriers 
```{r}
theme.del_dup_readthough_zscore <- function(p, base_size = 11, base_family = "Helvetica") {
  p <- p + theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, size = border_size),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_blank(),
          axis.ticks.x = element_line(size=ticks_size),
          axis.ticks.y = element_line(size=ticks_size),
          axis.ticks.length = unit(.15, "cm"),
          axis.title.x = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.title.y = element_text(size=base_size+2, family=base_family, face="plain", angle=90, margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.text.y = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 1, 0, 0), "mm"), color="black"),
          axis.text.x = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, margin = unit(c(1, 0, 0, 0), "mm"), color="black"),
          legend.position = "none", 
          legend.key = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          complete = TRUE,
          strip.placement = "outside",
          strip.background = element_blank(),
          strip.text = element_blank(), #element_text(size=base_size+2, family = base_family, face="plain", hjust=0.5, vjust=1),
          plot.title = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, vjust=2, color="black"),
          panel.spacing = unit(1, "lines")
          )
  return(p)
}

readthrough_rpkm_zscore_pdf = file.path(fig4_out_dir, "del_dup_tx_readthrough_fbnc_fpkm_zscore.pdf")
pdf(readthrough_rpkm_zscore_pdf, width = 4,height = 4, useDingbats=FALSE)

del_dup_rpkm_cov_zscore <- ggplot() +  
  rasterize(geom_point(data = tx_readthrough_noncarrier_df, aes(x=fpkm_zscore, y=cov_fraction_zscore, color=carrier, fill=carrier), alpha=0.3, size=4), dpi=600) +
  geom_point(data = tx_readthrough_carrier_df, aes(x=fpkm_zscore, y=cov_fraction_zscore, color=carrier, fill=carrier), alpha=0.8, size=4) + 
  scale_y_continuous(name="FBNC z-score") +
  scale_x_continuous(name="FPKM z-score") +
  scale_color_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication")) +
  scale_fill_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication"))
  
theme.del_dup_readthough_zscore(del_dup_rpkm_cov_zscore)
dev.off()
readthrough_rpkm_zscore_png <- file.path(fig4_out_dir, "del_dup_tx_readthrough_fbnc_fpkm_zscore.png")
theme.del_dup_readthough_zscore(del_dup_rpkm_cov_zscore)
ggsave(readthrough_rpkm_zscore_png, width=4, height=4, dpi=600)
```

### Relationship between misexpression level and readthrough region FPKM
```{r}
readthrough_rpkm_misexp_zscore_pdf = file.path(fig4_out_dir, "del_dup_tx_readthrough_fpkm_misexp_zscore.pdf")
pdf(readthrough_rpkm_misexp_zscore_pdf, width = 4,height = 4, useDingbats=FALSE)

del_dup_rpkm_misexp_zscore <- ggplot() +  
  rasterize(geom_point(data = tx_readthrough_noncarrier_df, aes(x=fpkm_zscore, y=z.score, color=carrier, fill=carrier), alpha=0.3, size=4), dpi=600) +
  geom_point(data = tx_readthrough_carrier_df, aes(x=fpkm_zscore, y=z.score, color=carrier, fill=carrier), alpha=0.8, size=4) + 
  scale_x_continuous(name="FPKM z-score") +
  scale_y_continuous(name="Misexpression z-score") +
  scale_color_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication")) +
  scale_fill_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication"))
  
theme.del_dup_readthough_zscore(del_dup_rpkm_misexp_zscore)
dev.off()
readthrough_rpkm_misexp_zscore_png <- file.path(fig4_out_dir, "del_dup_tx_readthrough_fpkm_misexp_zscore.png")
theme.del_dup_readthough_zscore(del_dup_rpkm_misexp_zscore)
ggsave(readthrough_rpkm_misexp_zscore_png, width=4, height=4, dpi=600)
```

### Relationship between readthrough FBNC and misexpression level 
```{r}
readthrough_cov_misexp_zscore_pdf = file.path(fig4_out_dir, "del_dup_tx_readthrough_fbnc_misexp_zscore.pdf")
pdf(readthrough_cov_misexp_zscore_pdf, width = 4,height = 4, useDingbats=FALSE)

del_dup_cov_misexp_zscore <- ggplot() +  
  rasterize(geom_point(data = tx_readthrough_noncarrier_df, aes(x=cov_fraction_zscore, y=z.score, color=carrier, fill=carrier), alpha=0.3, size=4), dpi=600) +
  geom_point(data = tx_readthrough_carrier_df, aes(x=cov_fraction_zscore, y=z.score, color=carrier, fill=carrier), alpha=0.8, size=4) +
  scale_x_continuous(name="FBNC z-score") +
  scale_y_continuous(name="Misexpression z-score") +
  scale_color_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication")) +
  scale_fill_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication"))
  
theme.del_dup_readthough_zscore(del_dup_cov_misexp_zscore)
dev.off()
readthrough_cov_misexp_zscore_png <- file.path(fig4_out_dir, "del_dup_tx_readthrough_fbnc_misexp_zscore.png")
theme.del_dup_readthough_zscore(del_dup_cov_misexp_zscore)
ggsave(readthrough_cov_misexp_zscore_png, width=4, height=4, dpi=600)
```

### Fig. Readthrough region length DEL and DUPs  
```{r}
theme.misexp_svlen <- function(p, base_size = 11, base_family = "Helvetica") {
  p <- p + theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, size = border_size),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_line(size=ticks_size),
          axis.ticks.length = unit(.15, "cm"),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size=base_size+2, family=base_family, face="plain", angle=90, margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.text.y = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 1, 0, 0), "mm"), color="black"),
          axis.text.x = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, margin = unit(c(1, 0, 0, 0), "mm"), color="black"),
          legend.key = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          legend.position = "none",
          complete = TRUE,
          strip.placement = "bottom",
          strip.background = element_blank(),
          strip.text = element_blank(), #element_text(size=base_size+2, family = base_family, face="plain", hjust=0.5, vjust=1),
          plot.title = element_text(hjust=0.5), 
          panel.spacing = unit(1, "lines")
          )
  return(p)
}

tx_readthrough_len_df <- tx_readthrough_carrier_df %>%
  select(vrnt_id, SVTYPE, total_len) %>%
  distinct()

readthrough_len_fig_path = file.path(fig4_out_dir, "readthrough_region_len.pdf")
pdf(readthrough_len_fig_path, width = 3.5,height = 4, useDingbats=FALSE)

set.seed(39)
readthrough_total_len <- ggplot(tx_readthrough_len_df, aes(x=SVTYPE, y=total_len,  color=SVTYPE, fill=SVTYPE)) + 
  geom_violin(position = position_dodge(0.9), alpha=0.2, color="white", width=1, bw=0.2) + 
  geom_jitter(width=0.2, fill="black", size=4) +
  scale_x_discrete(name="",
                   labels=c("Deletions", "Duplications")
                   ) +
  scale_y_continuous(name = "Readthrough region length",
                     trans="log10", 
                     breaks= c(100, 1000, 10000, 100000),
                     labels= c("100bp","1kb", "10kb", "100kb")) +
  scale_color_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication")) +
  scale_fill_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication"))

theme.misexp_svlen(readthrough_total_len)
dev.off()
readthrough_len_fig_png <- file.path(fig4_out_dir, "readthrough_region_len.png")
theme.misexp_svlen(readthrough_total_len)
ggsave(readthrough_len_fig_png, width=3.5, height=4, dpi=600)
```

### Fig. FPKM per variant-gene pair across all samples 
```{r}
### Supplementary raw FPKM per variant-gene pair
theme.fpkm_coverage_raw <- function(p, base_size = 11, base_family = "Helvetica") {
  p <- p + theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, size = border_size),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_blank(),
          axis.ticks.x = element_line(size=ticks_size),
          axis.ticks.y = element_line(size=ticks_size),
          axis.ticks.length = unit(.15, "cm"),
          axis.title.x = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.title.y = element_text(size=base_size+2, family=base_family, face="plain", angle=90, margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.text.y = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 1, 0, 0), "mm"), color="black"),
          axis.text.x = element_blank(),
          legend.position = "bottom", 
          legend.key = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          complete = TRUE,
          strip.placement = "bottom",
          strip.background = element_blank(),
          strip.text = element_blank(), #element_text(size=base_size+2, family = base_family, face="plain", hjust=0.5, vjust=1),
          plot.title = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, vjust=2, color="black"),
          panel.spacing = unit(1, "lines")
          )
  return(p)
}

readthrough_fpkm_raw_path = file.path(supp_fig13_out_dir, "readthrough_fpkm_raw.pdf")
pdf(readthrough_fpkm_raw_path, width = 6,height = 4, useDingbats=FALSE)

readthrough_fpkm_raw <- ggplot() +  
  rasterize(geom_jitter(data = tx_readthrough_noncarrier_df, aes(x=vrnt_gene_id, y=fpkm, color=carrier, fill=carrier), alpha=0.3, size=4, width=0.1), dpi=600) +
  geom_point(data = tx_readthrough_carrier_df, aes(x=vrnt_gene_id, y=fpkm, color=carrier, fill=carrier), alpha=0.8, size=4) + 
  scale_y_continuous(name="FPKM") +
  scale_x_discrete(name="Predicted readthrough region") +
  scale_color_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication")) +
  scale_fill_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication"))
  
theme.fpkm_coverage_raw(readthrough_fpkm_raw)
dev.off()
readthrough_fpkm_raw_png = file.path(supp_fig13_out_dir, "readthrough_fpkm_raw.png")
theme.fpkm_coverage_raw(readthrough_fpkm_raw)
ggsave(readthrough_fpkm_raw_png, width=6, height=4, dpi=600)
```

### Fig. FBNC per variant-gene pair across all samples 
```{r}
### Supplementary raw coverage per variant-gene pair
readthrough_coverage_raw_path = file.path(supp_fig13_out_dir, "readthrough_fbnc_raw.pdf")
pdf(readthrough_coverage_raw_path, width = 6,height = 4, useDingbats=FALSE)

readthrough_coverage_raw <- ggplot() +  
  rasterize(geom_jitter(data = tx_readthrough_noncarrier_df, aes(x=vrnt_gene_id, y=cov_fraction, color=carrier, fill=carrier), alpha=0.3, size=4, width=0.1), dpi=600) +
  geom_point(data = tx_readthrough_carrier_df, aes(x=vrnt_gene_id, y=cov_fraction, color=carrier, fill=carrier), alpha=0.8, size=4) + 
  scale_y_continuous(name="FBNC") +
  scale_x_discrete(name="Predicted readthrough region") +
  scale_color_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication")) +
  scale_fill_manual(name = "", values = sv_vrnt_colors,  labels=c("DEL"="Deletion", "DUP"= "Duplication"))
  
theme.fpkm_coverage_raw(readthrough_coverage_raw)
dev.off()
readthrough_coverage_raw_png = file.path(supp_fig13_out_dir, "readthrough_fbnc_raw.png")
theme.fpkm_coverage_raw(readthrough_coverage_raw)
ggsave(readthrough_coverage_raw_png, width=6, height=4, dpi=600)
```

### DEL & DEL readthrough VEP consequences 
```{r}
tx_readthrough_consq_path <-  file.path(wkdir, "6_misexp_dissect/tx_readthrough/combine/del_dup_tx_readthrough_consq.tsv") 
tx_readthrough_consq_df <- read.csv(tx_readthrough_consq_path, sep='\t')

theme.tx_readthrough_sv_types <- function(p, base_size = 11, base_family = "Helvetica") {
  p <- p + theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, size = border_size),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_line(size=ticks_size),
          axis.ticks.length = unit(.15, "cm"),
          axis.title.x = element_text(size=base_size+2, family=base_family, face="plain", vjust=-1, margin = unit(c(2, 0, 0, 0), "mm"), color="black"),
          axis.title.y = element_text(size=base_size+2, family=base_family, face="plain", angle=90, margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.text.y = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 1, 0, 0), "mm"), color="black"),
          axis.text.x = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, margin = unit(c(1, 0, 0, 0), "mm"), color="black"),
          legend.key = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          legend.position = "bottom",
          complete = TRUE,
          strip.placement = "outside",
          strip.background = element_blank(),
          strip.text = element_blank(), #element_text(size=base_size+2, family = base_family, face="plain", hjust=0.5, vjust=1),
          plot.title = element_text(hjust=0.5), 
          panel.spacing = unit(1, "lines")
          )
  return(p)
}
consq_order <-c("Transcript amplification", "Upstream (5 kb)", "Non-coding transcript","No predicted effect")

tx_readthrough_consq_df <- tx_readthrough_consq_df %>%
  mutate(
    consq_name = factor(consq_name, levels=consq_order),
  )

consq_colors <- c("Transcript amplification"="#76438C",
                 "Upstream (5 kb)"="#9681FF", 
                 "Non-coding transcript"="#A4A0CA", 
                 "No predicted effect"="#E2E2E2") 

tx_readthrough_del_dup_consq_path = file.path(supp_fig13_out_dir, "tx_readthrough_del_dup_consq.pdf")
pdf(tx_readthrough_del_dup_consq_path, width = 3.5,height = 5, useDingbats=FALSE)

tx_readthrough_vrnt_consq_fig <- ggplot(tx_readthrough_consq_df, aes(x=SVTYPE, y=vrnt_count, label=vrnt_count, color=consq_name, fill=consq_name)) +  
  geom_bar(width = 0.8, stat="identity", position="fill") +
  scale_x_discrete(name = "", 
                   expand=c(0,0), 
                   labels = c("DEL" = "Deletions", "DUP"="Duplications")
                   ) +
  scale_y_continuous(name = "Variant percentage", 
                     expand=c(0.001, 0.001), 
                     breaks=c(0, 0.25, 0.5, 0.75, 1.0),
                     labels=c("0%", "25%", "50%", "75%", "100%"),
                     ) + 
  facet_wrap(SVTYPE~., ncol=2, scales="free_x") + 
  geom_text(size = 5.5, position = position_fill(vjust = 0.5), color="white") + 
  scale_fill_manual(labels=consq_order, values=consq_colors, name="") +
  scale_color_manual(labels=consq_order, values=consq_colors, name="") + 
  guides(fill=guide_legend(ncol=1, bycol=TRUE)) 

theme.tx_readthrough_sv_types(tx_readthrough_vrnt_consq_fig)
dev.off()
tx_readthrough_del_dup_consq_png = file.path(supp_fig13_out_dir, "tx_readthrough_del_dup_consq.png")
theme.tx_readthrough_sv_types(tx_readthrough_vrnt_consq_fig)
ggsave(tx_readthrough_del_dup_consq_png, width=3.5, height=5, dpi=600)
```
### RTP1 misexpression carrier vs non-carriers
```{r}
### RTP1 misexpression level 
rtp1_carrier_path = file.path(wkdir, "6_misexp_dissect/examples/RTP1_DEL_chr3_187069321_187094542_carrier_noncarrier.tsv")
rtp1_carrier_df <- read.csv(rtp1_carrier_path, sep="\t")

ticks_size <- 0.75
theme.misexp_carrier_fig <- function(p, base_size = 11, base_family = "Helvetica") {
  p <- p + theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, size = border_size),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_line(size=ticks_size),
          axis.ticks.length = unit(.15, "cm"),
          axis.title.x = element_text(size=base_size+2, family=base_family, face="plain", vjust=-1, margin = unit(c(2, 0, 0, 0), "mm"), color="black"),
          axis.title.y = element_text(size=base_size+2, family=base_family, face="plain", angle=90, margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.text.y = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 1, 0, 0), "mm"), color="black"),
          axis.text.x = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, margin = unit(c(1, 0, 0, 0), "mm"), color="black"),
          legend.key = element_blank(),
          legend.title = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          legend.text = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          legend.position = "none", 
          complete = TRUE,
          strip.placement = "outside",
          strip.background = element_blank(),
          strip.text = element_text(size=base_size+2, family = base_family, face="plain", hjust=0.5),
          plot.title = element_text(hjust=0.5, size=base_size, family=base_family, face="bold", vjust=2), 
          plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
          )
  return(p)
}

rtp1_carrier_df$status <- factor(rtp1_carrier_df$status, levels = c("Non-carrier", "Carrier"))

misexp_color <- c("Misexpressed" ='#FF4848', # misexoression 
                   'Not misexpressed' = '#909090' # grey
                         )

rtp1_carrier_tpm_fig <- file.path(fig4_out_dir, "rtp1_carrier_noncarrier.pdf")
pdf(rtp1_carrier_tpm_fig, width = 3, height = 4, useDingbats=FALSE)

rtp1_carrier_tpm_fig <- ggplot(rtp1_carrier_df, aes(x=status, y=TPM, color=misexp)) + 
  rasterize(geom_point(position = position_jitter(width = 0.125, height = 0), size=3, alpha=0.75), dpi=600) +
  scale_x_discrete(name = "", labels=c("Non-carrier\n (0/0)", "Carrier\n (0/1)")) +
  scale_y_continuous(name = "Expression (TPM)",  expand=c(0.01,0.01), limits=c(0, 9), breaks=c(0, 2, 4, 6, 8)) + 
  ggtitle("RTP1") + 
  scale_color_manual(name = "", values = misexp_color) 

theme.misexp_carrier_fig(rtp1_carrier_tpm_fig)
dev.off()
rtp1_carrier_tpm_fig_png <- file.path(fig4_out_dir, "rtp1_carrier_noncarrier.png")
theme.misexp_carrier_fig(rtp1_carrier_tpm_fig)
ggsave(rtp1_carrier_tpm_fig_png, width=3, height=4, dpi=600)
```

### OTP misexpression carriers vs non-carriers
```{r}
otp_carrier_path = file.path(wkdir, "6_misexp_dissect/examples/OTP_DEL_chr5_77674588_77771600_carrier_noncarrier.tsv")
otp_carrier_df <- read.csv(otp_carrier_path, sep="\t")

ticks_size <- 0.75
theme.misexp_carrier_fig <- function(p, base_size = 11, base_family = "Helvetica") {
  p <- p + theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(),
          panel.border = element_rect(color = "black", fill = NA, size = border_size),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_line(size=ticks_size),
          axis.ticks.length = unit(.15, "cm"),
          axis.title.x = element_text(size=base_size+2, family=base_family, face="plain", vjust=-1, margin = unit(c(2, 0, 0, 0), "mm"), color="black"),
          axis.title.y = element_text(size=base_size+2, family=base_family, face="plain", angle=90, margin = unit(c(0, 2, 0, 0), "mm"), color="black"),
          axis.text.y = element_text(size=base_size+2, family=base_family, face="plain", margin = unit(c(0, 1, 0, 0), "mm"), color="black"),
          axis.text.x = element_text(size=base_size+2, family=base_family, face="plain", hjust=0.5, margin = unit(c(1, 0, 0, 0), "mm"), color="black"),
          legend.key = element_blank(),
          legend.title = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          legend.text = element_text(size=base_size+2, family = base_family, face="plain", color="black"),
          legend.position = "none", 
          complete = TRUE,
          strip.placement = "outside",
          strip.background = element_blank(),
          strip.text = element_text(size=base_size+2, family = base_family, face="plain", hjust=0.5),
          plot.title = element_text(hjust=0.5, size=base_size, family=base_family, face="bold", vjust=2), 
          plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
          )
  return(p)
}

otp_carrier_df$status <- factor(otp_carrier_df$status, levels = c("Non-carrier", "Carrier"))

misexp_color <- c("Misexpressed" ='#FF4848', # misexpression 
                   'Not misexpressed' = '#909090' # grey
                         )


otp_carrier_tpm_fig <- file.path(supp_fig13_out_dir, "otp_carrier_noncarrier.pdf")
pdf(otp_carrier_tpm_fig, width = 3, height = 4, useDingbats=FALSE)

otp_carrier_tpm_fig <- ggplot(otp_carrier_df, aes(x=status, y=TPM, color=misexp)) + 
  rasterize(geom_point(position = position_jitter(width = 0.125, height = 0), size=3, alpha=0.75)) +
  scale_x_discrete(name = "", labels=c("Non-carrier\n (0/0)", "Carrier\n (0/1)")) +
  scale_y_continuous(name = "Expression (TPM)",  expand=c(0.01,0.01), limits=c(0, 7), breaks=c(0, 2, 4, 6, 8)) + 
  ggtitle("OTP") + 
  scale_color_manual(name = "", values = misexp_color) 

theme.misexp_carrier_fig(otp_carrier_tpm_fig)
dev.off()
otp_carrier_tpm_fig_png <- file.path(supp_fig13_out_dir, "otp_carrier_noncarrier.png")
theme.misexp_carrier_fig(otp_carrier_tpm_fig)
ggsave(otp_carrier_tpm_fig_png, width=3, height=4, dpi=600)
```


